### 2011-09-30 proposal/paper.tex
Adding proposal template

\documentstyle[12pt,psfig]{article}

%set dimensions of columns, gap between columns, and space between paragraphs
\setlength{\textheight}{8.75in}
% \setlength{\columnsep}{2.0pc}
\setlength{\textwidth}{6.8in}
\setlength{\footheight}{0.0in}
\setlength{\topmargin}{0.25in}
\setlength{\headheight}{0.0in}
\setlength{\headsep}{0.0in}
\setlength{\oddsidemargin}{-.19in}
\setlength{\parindent}{1pc}
\renewcommand{\baselinestretch}{1.3}

% change the section numbering format:
% \renewcommand{\thesection}{\Alph{section}}

\begin{document}

% \psfigurepath{FIG}

\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\bit}{\begin{itemize}}
\newcommand{\eit}{\end{itemize}}
\newcommand{\goal}[1]{ {\noindent {$\Rightarrow$} \em {#1} } }
\newcommand{\hide}[1]{}
\newcommand{\comment}[1]{ {\footnotesize {#1} } }
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{proof}{Proof}
\newtheorem{defn}{Definition}
\newtheorem{algo}{Algorithm}
\newtheorem{observation}{Observation}

% print DRAFT across each ps page
% \special{!userdict begin /bop-hook{gsave 200 30 translate
% 65 rotate /Times-Roman findfont 216 scalefont setfont 0 0 moveto 0.95
% setgray (DRAFT) show grestore}def end}

\title{Catchy Title}

\author{ {\em John Smith} \\
	    Dept. of Computer Science \\
	    Carnegie Mellon University \\
	    {\tt jsmith@cs.cmu.edu}
	 \and
	 {\em Tom Johnson} \\
	     Dept. of Metaphysics \\
	     Confusion State University \\
	     {\tt tj@bla.csu.edu}
        }

\maketitle
\begin{abstract}
We solved an important problem,
outperforming all previous attempts.
\end{abstract}

\section{Introduction}
\label{sec:intro}
\bit
\item
what is the problem
\item
what are the applications
\eit

\section{Survey}
\label{sec:survey}
All the solutions that
have been proposed in the past!
Example of citation: \cite{Aho75Fast}

Table \ref{tab:symbols} gives a list of symbols used.
\begin{table}[htb]
\begin{center} 
\begin{tabular}{|l | c | } \hline \hline 
Symbol & Definition \\ \hline
$\pi$ & 3.14 \\
$f$  & frequency \\ \hline
\end{tabular} 
\end{center} 
\caption{Symbol table}
\label{tab:symbols} 
 \end{table}

\section{Proposed Method}
\label{sec:proposed}
Our proposed method is as follows: ...

\section{Experiments}
\label{sec:experiments}
We implemented our method
and compared it with the older ones.
The results are very promising.

Figure \ref{fig:results} shows the results
(and, actually, how to insert postscript files):
\begin{figure}[htbf]
\begin{center}
\begin{tabular}{cc}
     % uncomment the next lines, and give the right ps files
     \psfig{figure=figs/data.ps,width=2in} &
     \psfig{figure=figs/plot.ps,width=2in} \\
    (a) & (b) 
\end{tabular}
\caption{A fictitious dataset (a) and its performance plot (b)}
\label{fig:results}
\end{center}
\end{figure}

\section{Conclusions}
\label{sec:conclusions}
The proposed method outperforms older approaches
in speed and quality.

\bibliography{christosref,other}
\bibliographystyle{plain}
\newpage
\pagenumbering{roman}
\tableofcontents

\end{document}